Remote sensing the gas transfer velocity using SWOT observations

### 2020-12-07 Outline.Rmd
Figures update

## Goal 1
#### Theory
1) Raymond 2012 show that k600 can be scaled using hydraulics b/c k600 generally correlates with the turbulent dissipation rate (eD = gSV)
2) Ulseth et al 2019 show two distinct scaling regimes for k600 with eD using an expanded dataset including steeper streams than those in Raymond 2012.
3) Here, we take the Ulseth et al. 2019 dataset and develop a k600~eD scaling function that accounts for this regime shift and the fact that scaling relationships differ as a f(river size). BUT, we only use SWOT observables to determine which scaling relationship to use.
    a. i.e. river width and slope alone must be used to choose which scaling model to use.
	    i. $k_{600}=a(gSV)^{b}$ where a and b are differentially determined by river size.
	  b. Small rivers had to be broken into high/low slope to maintain all linear regression assumptions.

![**Figure 1** *eD~k600 scaling by river width/slope. This is how we are able to scale k600 using just SWOT observables*](outputs//k600//width_eD_theory.jpg)

4) To test this, we then combine these five scaling functions into a single 'rule-based' regression model, train it on 80% of the data and validate it on the remaining 20% (below). The version ultimately implemented in our algorithm is trained on the entire dataset however.

![**Figure 2** *Model validation on 20% of data, withheld from training*](outputs//k600//validation.jpg)

@@

![**Figure 5** *Pepsi validation by river*](outputs//validation//validation_by_river.jpg)

5. Below are timeseries plots of observed versus modeled k600 for three rivers randomly sampled from the three tertiles of the by-river kge scores. a) is a 'good' river, b is a 'reasonabley performing' river, and c is a 'bad' performing river.

![**Figure 6** *Pepsi validation timeseries for three rivers, where the model results include the posterior means and 95% CIs: a) randomly selected from upper tertile of kge scores b) randomly selected from middle tertile, c) randomly selected from worse tertile*](outputs//validation//validation_timeseries.jpg)

>>>>

## Goal 1
#### Theory
1) Raymond 2012 show that k600 can be scaled using hydraulics b/c k600 generally correlates with the turbulent dissipation rate (eD = gSV)
2) Ulseth et al 2019 show two distinct scaling regimes for k600 with eD using an expanded dataset including steeper streams than those in Raymond 2012.
3) Here, we take the Ulseth et al. 2019 dataset and develop a k600~eD scaling function that accounts for this regime shift and the fact that scaling relationships differ as a f(river size). BUT, we only use SWOT observables to determine which scaling relationship to use.
    a. i.e. river width and slope alone must be used to choose which scaling model to use.
	    i. $k_{600}=a(gSV)^{b}$ where a and b are differentially determined by river size.
	  b. Small rivers had to be broken into high/low slope to maintain all linear regression assumptions.
4) To test this, we then combine these five scaling functions into a single 'rule-based' regression model, train it on 80% of the data and validate it on the remaining 20% (below). The version ultimately implemented in our algorithm is trained on the entire dataset however.

![**Figure 2** *a) eD~k600 scaling by river width/slope. This is how we are able to scale k600 using just SWOT observables. b) Overall model validation on 20% of data, withheld from training*](outputs//k600//validation.jpg)

@@

![**Figure 5** *a) Performance metrics by river. b)-d): validation timeseries for three rivers, where the model results include the posterior means and 95% CIs: b) randomly selected from upper tertile of kge scores c) randomly selected from middle tertile, d) randomly selected from worse tertile*](outputs//validation//validation_by_river.jpg)
